IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-07 paper/appendix.tex
Finish proof of helper lemma.

\begin{lemma}
  \label{lemma:high-not-blocking}
  For any task |it|, task lists |its|, store |iS|, and label |il|, if
  $|erase il it|=|bullet|$, then there exists a task list
  |its'| and a store |iS'| such that
  \[
    |iconf iS (it, its)| |.->| |iconf iS' (its, its')|
    \qquad
    |erase il its'|=|nil|
    \qquad
    |erase il iS'|=|erase il iS|
  \]
\end{lemma}
\begin{proof}
  TODO (consideration of all rules that could be applied)
\end{proof}

@@

    Now, the first task |it2| in |ic2''| is |il|-equivalent to the task |it1|.
    This implies that the two tasks must have the same id, label and
    can only differ in the expression or store if some subexpression
    is of the form |Labeled il' ie|.  In this case, the expression |ie| could
    be different in the two threads if |lcurr canFlowTo il'|.  However, none of the reduction rules
    depend on an expression in that position, and there is never a
    hole in that position
    where evaluation could take place.  Thus, the same rules will syntactically
    match for both task, and we are left to argue that all premises
    evaluate to the same values for |it1| and |it2|, as well as that
    the resulting states |iS1'| and
    |iS2''| are |il|-equivalent.
    The additional premises $P$ that follow
    the condition in Definition~\ref{def:restricted} are not a problem,
    since those
    predicates only depend on |erase il ic1|, which is equivalent
    to |erase il ic2''|, and thus those predicates evaluate in the same way.
    All other premises are either on the threads labels (which are the same),
    or on the state |iS1|, or |iS2'|, respectively.  Because
    |erase il iS1=erase il iS2'|, all of these also evaluate in the same way,
    as can be seen by simply considering all rules that involve or
    change the state:
    \begin{description}
      \item[Case \textsc{I-send}]
      Here, the task |it2| will send the same message to the same
      receiver queue. This
      queue is either completely erased, or it is |il|-equivalent.  In both
      cases, |il|-equivalence of |iS1'| and |iS2'| is preserved.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      When the tasks are receiving a message, then by the reduction rules
      we know that they first filter the queue by the label
      |lcurr| of |it1|.  We
      also know that the queues are equivalent when filtered by the less
      restrictive label |il|, thus the messages received (or dropped) from the
      queue are equivalent.
      \item[Case \textsc{I-new}] The newly allocated address can be the same
      for both |it1| and |it2|, thus resulting in |il|-equivalent states.
      \item[Case \textsc{I-write}] By |erase il it1=earse il it2| both tasks
      write the same value, and therefore the resulting states will still
      be |il|-equivalent.
    \end{description}

  \end{itemize}
\end{proof}

>>>>

\begin{lemma}
  \label{lemma:high-not-blocking}
  For any task |it|, task lists |its|, store |iS|, and label |il|, if
  $|erase il it|=|bullet|$, then there exists a task list
  |its'| and a store |iS'| such that
  \begin{align}
  |iconf iS (it, its)| |.->| |iconf iS' (its, its')| \label{eq:hnb-1} \\
  |erase il its'|=|nil| \label{eq:hnb-2}\\
  |erase il iS'|=|erase il iS| \label{eq:hnb-3}
  \end{align}
\end{lemma}
\begin{proof}
  From $|erase il it|=|bullet|$ we know that the current label |lcurr| of |it|
  must be above |il|.  Furthermore, tasks can always take a step (if no
  regular rule applies, then \textsc{I-noStep} can be used), and thus
  we consider all rules that could be applied to execute |it|.
  \begin{description}
    \item[Case \textsc{I-noStep} and \textsc{I-done}]
    In this case, the task |it| is dropped,
    and thus |its'=nil| and |iS'=iS| satisfy
    conditions~\eqref{eq:hnb-2} and~\eqref{eq:hnb-3}.
    \item[Case \textsc{I-sandbox}]
    The newly created task has a label of at least |lcurr|, and will thus be
    erased, as required by condition~\eqref{eq:hnb-2}.  Furthermore, the
    state only changes for the newly created thread, and thus the state
    change is erased, showing~\eqref{eq:hnb-3}.
  \end{description}
  In all other rules, no new tasks are created, and thus |its'| consists of just
  the one task |it'|, to which |it| executed.  Since the tasks label can
  only increase, |it'| is still erased, showing condition~\eqref{eq:hnb-2}.
  We are left to show condition~\eqref{eq:hnb-3} for the remaining rules.
  \begin{description}
    \item[Case \textsc{I-send}]
    A new message triple with label |il'| gets added to the message
    queue of the receiving thread.  However, since |lcurr canFlowTo il'|,
    the triple will get erased.
    \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
    In this case, only the queue of
    task |it| can change, which gets erased.
    \item[Case \textsc{I-new}] The newly allocated address has to be at a
    label at least as high as |lcurr|, and will thus be erased.
    \item[Case \textsc{I-write}] Only addresses with a label |il'| above
    |lcurr| can be written, thus the change in |iS1| will get erased.
    \item[Otherwise.]  None of the other rules modify the state |iS|, and
    thus |iS'=iS| will trivially satisfy condition~\eqref{eq:hnb-3}.
  \end{description}
\end{proof}

@@

    Now, the first task |it2| in |ic2''| is |il|-equivalent to the task |it1|.
    This implies that the two tasks must have the same id, label and
    can only differ in the expression or store if some subexpression
    is of the form |Labeled il' ie|.  In this case, the expression |ie| could
    be different in the two threads if |lcurr canFlowTo il'|.  However, none of the reduction rules
    depend on an expression in that position, and there is never a
    hole in that position
    where evaluation could take place.  Thus, the same rules will syntactically
    match for both task, and we are left to argue that all premises
    evaluate to the same values for |it1| and |it2|, as well as that
    the resulting states |iS1'| and
    |iS2''| are |il|-equivalent.
    The additional premises $P$ that follow
    the condition in Definition~\ref{def:restricted} are not a problem,
    since those
    predicates only depend on |erase il ic1|, which is equivalent
    to |erase il ic2''|, and thus those predicates evaluate in the same way.
    All other premises are either on the threads labels (which are the same),
    or on the state |iS1|, or |iS2'|, respectively.  Because
    |erase il iS1=erase il iS2'|, all of these also evaluate in the same way,
    as can be seen by simply considering all rules that involve or
    change the state:
    \begin{description}
      \item[Case \textsc{I-send}]
      Here, the task |it2| will send the same message to the same
      receiver queue. This
      queue is either completely erased, or it is |il|-equivalent.  In both
      cases, |il|-equivalence of |iS1'| and |iS2'| is preserved.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      When the tasks are receiving a message, then by the reduction rules
      we know that they first filter the queue by the label
      |lcurr| of |it1|.  We
      also know that the queues are equivalent when filtered by the less
      restrictive label |il|, thus the messages received (or dropped) from the
      queue are equivalent.
      \item[Case \textsc{I-new}] The newly allocated address can be the same
      for both |it1| and |it2|, thus resulting in |il|-equivalent states.
      \item[Case \textsc{I-write}] By |erase il it1=earse il it2| both tasks
      write the same value, and therefore the resulting states will still
      be |il|-equivalent.
    \end{description}
    After |it2| has taken a step, we finally arrive in the desired
    configuration |ic2'=iconf iS2'' (its2'', its2''', its2'''')|, where
    |its2''''| contains the task resulting from executing |it2| (and
    might contain, zero (if the task was done or stuck), one (for most steps) or two tasks if a new task was launched).
    As required, we have
    \[ |ic2| |.->|^* |ic2''| |.->| |ic2'|
    \quad \land \quad |ic1'| \approx_{|il|} |ic2'| \text{.}\]
  \end{itemize}
\end{proof}
